---
title: "Plotting complex multimorbidity"
output: html_document
date: '2022-05-09'
---
## Load packages

```{r setup, include=FALSE}
library(tidyverse)
library(epitools)
library(ggsci) # for lancet colour scale
library(ggrepel)
library(data.table)
library(ggforce)
library(ggthemes)

theme_set(theme_minimal(base_size = 20) +
            theme(panel.grid.minor = element_blank(),
                  panel.grid.major.x = element_blank(),
                  panel.grid.major.y = element_line(size = 0.25),
                  axis.line = element_line(colour = "black"),
                  axis.title.y = element_text(margin = margin(0, 20, 0, 0)),
                  axis.title.x = element_text(margin = margin(20, 0, 0, 0)),
                  plot.caption = element_text(colour = "#AAAAAA"),
                  plot.margin = margin(3, 15, 3, 3, "mm")))
```

## Load dataset


```{r cars}
table_join_cmm <- read_csv("table_cmm_join.csv") 
```

## All definitions plot - prep dataframe

```{r pressure, echo=FALSE}
## Age and definition plot
colnames(table_join_cmm) <- c("age_gender_IMD", "Simple", "Complex", "Complex body system", "Complex physical mental", "n")

## One plot for both genders
df_plot <- table_join_cmm %>%
  pivot_longer(cols = c("Simple", "Complex", "Complex body system", "Complex physical mental"), names_to = "definition", values_to = "count") %>%
  replace(is.na(.),0) %>%
  separate("age_gender_IMD", c("age", "gender", "IMD")) %>%
  select(-IMD) %>%
  mutate(age = case_when(
   age == '00to04' ~ "00 - 04", 
   age == '05to09' ~ "05 - 09",
   age == '10to14' ~ "10 - 14", 
   age == '15to19' ~ "15 - 19",
   age == '20to24' ~ "20 - 24", 
   age == '25to29' ~ "25 - 29",
   age == '30to34' ~ "30 - 34", 
   age == '35to39' ~ "35 - 39",
   age == '40to44' ~ "40 - 44", 
   age == '45to49' ~ "45 - 49",
   age == '50to54' ~ "50 - 54", 
   age == '55to59' ~ "55 - 59",
   age == '60to64' ~ "60 - 64", 
   age == '65to69' ~ "65 - 69",
   age == '70to74' ~ "70 - 74", 
   age == '75to79' ~ "75 - 79",
   age == '80to84' ~ "80 - 84", 
   age == '85to89' ~ "85 - 89",
   age == 'Over90' ~ "90+", 
)) %>%
  group_by(age, definition) %>%
  summarise_if(is.numeric, sum) %>%
  mutate(prop = count/n *100)

df_plot$definition <- factor(df_plot$definition, levels = rev(c("Simple", "Complex", "Complex body system", "Complex physical mental")))

## One plot for each sex
df_plot_gender <- table_join_cmm %>%
  pivot_longer(cols = c("Simple", "Complex", "Complex body system", "Complex physical mental"), names_to = "definition", values_to = "count") %>%
  replace(is.na(.),0) %>%
  separate("age_gender_IMD", c("age", "gender", "IMD")) %>%
  select(-IMD) %>%
 mutate(age = case_when(
   age == '00to04' ~ "00 - 04", 
   age == '05to09' ~ "05 - 09",
   age == '10to14' ~ "10 - 14", 
   age == '15to19' ~ "15 - 19",
   age == '20to24' ~ "20 - 24", 
   age == '25to29' ~ "25 - 29",
   age == '30to34' ~ "30 - 34", 
   age == '35to39' ~ "35 - 39",
   age == '40to44' ~ "40 - 44", 
   age == '45to49' ~ "45 - 49",
   age == '50to54' ~ "50 - 54", 
   age == '55to59' ~ "55 - 59",
   age == '60to64' ~ "60 - 64", 
   age == '65to69' ~ "65 - 69",
   age == '70to74' ~ "70 - 74", 
   age == '75to79' ~ "75 - 79",
   age == '80to84' ~ "80 - 84", 
   age == '85to89' ~ "85 - 89",
   age == 'Over90' ~ "90+", 
)) %>%
   mutate(gender = case_when(
    gender == '1' ~ "Men",
    gender == '2' ~ "Women")) %>%
  group_by(age, gender, definition) %>%
  summarise_if(is.numeric, sum) %>%
  mutate(prop = count/n *100)

df_plot_gender$definition <- factor(df_plot_gender$definition, levels = c("Simple", "Complex", "Complex body system", "Complex physical mental"))
```

## All definitions plot code

```{r}
df_plot %>% 
  ggplot() +
  aes(age, prop, group = definition, colour = definition)+
  geom_line(size = 2)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = c(.2, .7)) +
  theme(legend.title=element_blank()) +
  scale_y_continuous(limits= c(0, 100), expand = c(0,0), 
                     breaks = scales::pretty_breaks(n = 4))+
  scale_x_discrete(expand = c(0,0))+
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "Age group (years)",
       y = "% population") +
  scale_colour_ptol()
  
ggsave("fig1_definitions_all.png", width = 16, height = 12)
ggsave("fig1_definitions_all.pdf", width = 16, height = 12)
```

## All definitions plot - by gender

```{r}
df_plot_gender %>% 
  ggplot() +
  aes(age, prop, group = gender, colour = gender)+
  geom_line(size = 2)+
  facet_wrap(~definition, scales="free")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = c(.1, .87)) +
  theme(legend.title=element_blank()) +
  scale_y_continuous(limits= c(0, 100), expand = c(0,0), 
                     breaks = scales::pretty_breaks(n = 4))+
  scale_x_discrete(expand = c(0,0))+
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "Age group (years)",
       y = "% population") +
  scale_colour_ptol()
  
ggsave("suppfig1_definitions_all.png", width = 16, height = 12)
ggsave("suppfig1_definitions_all.pdf", width = 16, height = 12)
```

## All definitions by IMD plot - prep dataframe

```{r}
df_plot_IMD <- table_join_cmm %>%
  pivot_longer(cols = c("Simple", "Complex", "Complex body system", "Complex physical mental"), names_to = "definition", values_to = "count") %>%
  replace(is.na(.),0) %>%
  separate("age_gender_IMD", c("age", "gender", "IMD Decile")) %>%
  select(-gender) %>%
  mutate(age = case_when(
   age == '00to04' ~ "00 - 04", 
   age == '05to09' ~ "05 - 09",
   age == '10to14' ~ "10 - 14", 
   age == '15to19' ~ "15 - 19",
   age == '20to24' ~ "20 - 24", 
   age == '25to29' ~ "25 - 29",
   age == '30to34' ~ "30 - 34", 
   age == '35to39' ~ "35 - 39",
   age == '40to44' ~ "40 - 44", 
   age == '45to49' ~ "45 - 49",
   age == '50to54' ~ "50 - 54", 
   age == '55to59' ~ "55 - 59",
   age == '60to64' ~ "60 - 64", 
   age == '65to69' ~ "65 - 69",
   age == '70to74' ~ "70 - 74", 
   age == '75to79' ~ "75 - 79",
   age == '80to84' ~ "80 - 84", 
   age == '85to89' ~ "85 - 89",
   age == 'Over90' ~ "90+", 
)) %>%
  mutate(`IMD Decile` = case_when(
    `IMD Decile` == '01' ~ "1 (least deprived)",
    `IMD Decile` == '02' ~ "2",
    `IMD Decile` == '03' ~ "3",
    `IMD Decile` == '04' ~ "4",
    `IMD Decile` == '05' ~ "5",
    `IMD Decile` == '06' ~ "6",
    `IMD Decile` == '07' ~ "7",
    `IMD Decile` == '08' ~ "8",
    `IMD Decile` == '09' ~ "9",
    `IMD Decile` == '10' ~ "10 (most deprived)"
  )) %>%
  group_by(age, `IMD Decile`, definition) %>%
  summarise_if(is.numeric, sum) %>%
  mutate(prop = count/n *100)

df_plot_IMD$definition <- factor(df_plot_IMD$definition, levels = c("Simple", "Complex", "Complex body system", "Complex physical mental"))

df_plot_IMD$`IMD Decile` <- factor(df_plot_IMD$`IMD Decile`, levels = c("1 (least deprived)", "2", "3", "4", "5", "6", "7", "8", "9", "10 (most deprived)"))
```

## All definitions plot by IMD - plot code

```{r}
df_plot_IMD %>% 
  ggplot() +
  aes(age, prop, group = `IMD Decile`, colour = `IMD Decile`)+
  geom_line(size = 1)+
  facet_wrap(~definition, scales="free")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = c(.1, .87)) +
  #theme(legend.title=element_blank()) +
  scale_y_continuous(limits= c(0, 100), expand = c(0,0), 
                     breaks = scales::pretty_breaks(n = 4))+
  scale_x_discrete(expand = c(0,0))+
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "Age group (years)",
       y = "% population") +
  scale_colour_ptol()
  
ggsave("suppfig2_definitions_IMD.png", width = 16, height = 12)
ggsave("suppfig2_definitions_IMD.pdf", width = 16, height = 12)
```

## All definitions IMD 1 and 10 only dataframe and plot code

```{r}
df_plot_IMD <- df_plot_IMD %>%
  filter(`IMD Decile` %in% c("1 (least deprived)", "10 (most deprived)")) %>%
  #filter(definition %in% c("Complex multimorbidity; physical-mental")) %>%
  rowwise %>%
  mutate(prop = ((prop.test(count, n, conf.level=0.95)$estimate)*100),
         lci = ((prop.test(count, n, conf.level=0.95)$conf.int[1])*100),
         uci = ((prop.test(count, n, conf.level=0.95)$conf.int[2])*100)) %>%
  mutate_if(is.numeric,
            round,
            digits = 2)

df_plot_IMD %>% 
  ggplot() +
  aes(age, prop, group = `IMD Decile`, colour = `IMD Decile`)+
  geom_line(size = 1)+
  geom_errorbar(aes(ymin=lci, ymax=uci), width=.1) +
  facet_wrap(~definition, scales="free")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = c(.1, .87)) +
  #theme(legend.title=element_blank()) +
  scale_y_continuous(limits= c(0, 100), expand = c(0,0), 
                     breaks = scales::pretty_breaks(n = 4))+
  scale_x_discrete(expand = c(0,0))+
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "Age group (years)",
       y = "% population") +
  scale_colour_ptol()
  
ggsave("fig2_definitions_IMD.png", width = 16, height = 12)
ggsave("fig2_definitions_IMD.pdf", width = 16, height = 12)
```

## Physical mental IMD 1 and 10 only dataframe and plot code (plus annotations)

```{r}
df_plot_mp_IMD <- df_plot_IMD %>%
  filter(`IMD Decile` %in% c("1 (least deprived)", "10 (most deprived)")) %>%
  filter(definition == "Complex physical mental")

df_plot_mp_IMD %>% 
  ggplot() +
  aes(age, prop, group = `IMD Decile`, colour = `IMD Decile`)+
  geom_line(size = 1)+
  geom_errorbar(aes(ymin=lci, ymax=uci), width=.1) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = c(.1, .87)) +
  scale_y_continuous(limits= c(0, 55), expand = c(0,0), 
                     breaks = scales::pretty_breaks(n = 4))+
  scale_x_discrete(expand = c(0,0))+
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "Age group (years)",
       y = "% population") +
  scale_colour_ptol() +
  annotate("segment", x = 10.5, xend = 19.5, y=37.18, yend = 37.18, colour = "black", linetype = 2)+
  annotate("segment", x = 9.5, xend = 19, y=33.19, yend = 33.19, colour = "black", linetype = 2)+
  annotate("text", x = 8.5, y = 38.8, label = "Rate in most deprived exceeds \n least deprived at any age", size = 5) +
  annotate("text", x = 7.5, y = 33, label = "Rate in most deprived > 40 years \nearlier than least deprived", size = 5)

ggsave("fig3_mp_IMD.png", width = 16, height = 12)
ggsave("fig3_mp_IMD.pdf", width = 16, height = 12)
```

## Physical mental plot - dataframe prep

```{r}
df_plot_mp_gender <- table_join_mp %>%
  pivot_longer(cols = c("Depression anxiety physical", "Dementia physical", "Substance alcohol misuse physical", "Schizophrenia bipolar physical"), names_to = "definition", values_to = "count") %>%
  replace(is.na(.),0) %>%
  separate("age_gender_IMD", c("age", "gender", "IMD")) %>%
  select(-IMD) %>%
 mutate(age = case_when(
    age == '00to09' ~ "0 - 9",
    age == '10to19' ~ "10 - 19",
    age == '20to29' ~ "20 - 29",
    age == '30to39' ~ "30 - 39",
    age == '40to49' ~ "40 - 49",
    age == '50to59' ~ "50 - 59",
    age == '60to69' ~ "60 - 69",
    age == '70to79' ~ "70 - 79",
    age == '80' ~ "80+"
  )) %>%
   mutate(gender = case_when(
    gender == '1' ~ "Men",
    gender == '2' ~ "Women")) %>%
  group_by(age, gender, definition) %>%
  summarise_if(is.numeric, sum) %>%
  mutate(prop = count/n *100)

df_plot_mp_gender$definition <- factor(df_plot_mp_gender$definition, levels = c("Depression anxiety physical", "Substance alcohol misuse physical", "Schizophrenia bipolar physical", "Dementia physical"))
```

## Physical mental plot code
## NOTE MUST CHANGE AGE BINS WHEN RELEASED FROM HIC

```{r}
df_plot_mp_gender %>% 
  ggplot() +
  aes(age, prop, group = gender, colour = gender)+
  geom_line(size = 2)+
  facet_wrap(~definition, scales="free")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = c(.1, .87)) +
  theme(legend.title=element_blank()) +
  scale_y_continuous(limits= c(0, 40), expand = c(0,0), 
                     breaks = scales::pretty_breaks(n = 4))+
  scale_x_discrete(expand = c(0,0))+
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "Age group (years)",
       y = "% population") +
  scale_colour_ptol()
  
ggsave("suppfig3_mpdefinitions_sex.png", width = 16, height = 12)
ggsave("suppfig3_mpdefinitions_sex.pdf", width = 16, height = 12)
```
