---
title: "Chapter 3 - definition of multimorbidity"
output: html_document
date: '2022-05-09'
---

## Load packages
```{r setup, include=FALSE}
library(tidyverse)
library(epitools)
library(ggsci) # for lancet colour scale
library(ggrepel)
library(data.table)
library(ggforce)
library(ggthemes)
library(scales)
install.packages("viridis")
library(viridis)

theme_set(theme_minimal(base_size = 20) +
            theme(panel.grid.minor = element_blank(),
                  panel.grid.major.x = element_blank(),
                  panel.grid.major.y = element_line(size = 0.25),
                  axis.line = element_line(colour = "black"),
                  axis.title.y = element_text(margin = margin(0, 20, 0, 0)),
                  axis.title.x = element_text(margin = margin(20, 0, 0, 0)),
                  plot.caption = element_text(colour = "#AAAAAA"),
                  plot.margin = margin(3, 15, 3, 3, "mm")))
```

## Load dataset
```{r cars}
table_cmm_join <- read_csv("table_cmm_join.csv") 
table_mp_join <- read_csv("table_mp_join.csv")
```

## Figure 1 - all definitions, x=age 
## Prep dataframe
```{r pressure, echo=FALSE}
# Change col names to plot labels
colnames(table_cmm_join) <- c("age_gender_IMD", "Multimorbidity 2+", "Multimorbidity 3+", "Multimorbidity 3+ from 3+", "Mental-physical multimorbidity", "n")

# Pivot and set age-bins
df_plot <- table_cmm_join %>%
  pivot_longer(cols = c("Multimorbidity 2+", "Multimorbidity 3+", "Multimorbidity 3+ from 3+", "Mental-physical multimorbidity"), names_to = "definition", values_to = "count") %>%
  replace(is.na(.),0) %>%
  separate("age_gender_IMD", c("age", "gender", "IMD")) %>%
  select(-IMD) %>%
  mutate(age = case_when(
   age == '00to04' ~ "00 - 04", 
   age == '05to09' ~ "05 - 09",
   age == '10to14' ~ "10 - 14", 
   age == '15to19' ~ "15 - 19",
   age == '20to24' ~ "20 - 24", 
   age == '25to29' ~ "25 - 29",
   age == '30to34' ~ "30 - 34", 
   age == '35to39' ~ "35 - 39",
   age == '40to44' ~ "40 - 44", 
   age == '45to49' ~ "45 - 49",
   age == '50to54' ~ "50 - 54", 
   age == '55to59' ~ "55 - 59",
   age == '60to64' ~ "60 - 64", 
   age == '65to69' ~ "65 - 69",
   age == '70to74' ~ "70 - 74", 
   age == '75to79' ~ "75 - 79",
   age == '80to84' ~ "80 - 84", 
   age == '85to89' ~ "85 - 89",
   age == 'Over90' ~ "90+", 
)) %>%
  group_by(age, definition) %>%
  summarise_if(is.numeric, sum) %>%
  mutate(prop = count/n *100)

# Factorise and set order
df_plot$definition <- factor(df_plot$definition, levels = rev(c("age_gender_IMD", "Multimorbidity 2+", "Multimorbidity 3+", "Multimorbidity 3+ from 3+", "Mental-physical multimorbidity")))

## Including gender stratification
df_plot_gender <- table_cmm_join %>%
  pivot_longer(cols = c("Multimorbidity 2+", "Multimorbidity 3+", "Multimorbidity 3+ from 3+", "Mental-physical multimorbidity"), names_to = "definition", values_to = "count") %>%
  replace(is.na(.),0) %>%
  separate("age_gender_IMD", c("age", "gender", "IMD")) %>%
  select(-IMD) %>%
 mutate(age = case_when(
   age == '00to04' ~ "00 - 04", 
   age == '05to09' ~ "05 - 09",
   age == '10to14' ~ "10 - 14", 
   age == '15to19' ~ "15 - 19",
   age == '20to24' ~ "20 - 24", 
   age == '25to29' ~ "25 - 29",
   age == '30to34' ~ "30 - 34", 
   age == '35to39' ~ "35 - 39",
   age == '40to44' ~ "40 - 44", 
   age == '45to49' ~ "45 - 49",
   age == '50to54' ~ "50 - 54", 
   age == '55to59' ~ "55 - 59",
   age == '60to64' ~ "60 - 64", 
   age == '65to69' ~ "65 - 69",
   age == '70to74' ~ "70 - 74", 
   age == '75to79' ~ "75 - 79",
   age == '80to84' ~ "80 - 84", 
   age == '85to89' ~ "85 - 89",
   age == 'Over90' ~ "90+", 
)) %>%
   mutate(gender = case_when(
    gender == '1' ~ "Men",
    gender == '2' ~ "Women")) %>%
  group_by(age, gender, definition) %>%
  summarise_if(is.numeric, sum) %>%
  mutate(prop = count/n *100)

# Factorise and set order
df_plot_gender$definition <- factor(df_plot_gender$definition, levels = c("age_gender_IMD", "Multimorbidity 2+", "Multimorbidity 3+", "Multimorbidity 3+ from 3+", "Mental-physical multimorbidity"))
```


## Figure 1 - all definitions, x=age 
```{r}
df_plot %>% 
  ggplot() +
  aes(age, prop, group = definition, colour = definition)+
  geom_line(size = 2)+
  theme(axis.ticks.x = element_line()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = c(.2, .7)) +
  theme(legend.title=element_blank()) +
  scale_y_continuous(limits= c(0, 100), expand = c(0,0), 
                     breaks = scales::pretty_breaks(n = 4))+
  scale_x_discrete(expand = c(0,0))+
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "Age group (years)",
       y = "% population") +
  scale_colour_ptol()
  
ggsave("fig1_definitions.png", width = 16, height = 12)
ggsave("fig1_definitions.pdf", width = 16, height = 12)
```

## Exclude or need to age-statndardise this
## Figure 1 - all definitions, x=age, stratified=gender
```{r}
df_plot_gender %>% 
  ggplot() +
  aes(age, prop, group = gender, colour = gender)+
  geom_line(size = 2)+
  facet_wrap(~definition, scales="free")+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))+
  theme(legend.position = c(.1, .87)) +
  theme(legend.title=element_blank()) +
  scale_y_continuous(limits= c(0, 100), expand = c(0,0), 
                     breaks = scales::pretty_breaks(n = 4))+
  scale_x_discrete(expand = c(0,0))+
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "Age group (years)",
       y = "% population") +
  scale_colour_ptol()
  
ggsave("suppfig1_definitions_gender.png", width = 16, height = 12)
ggsave("suppfig1_definitions_gender.pdf", width = 16, height = 12)
```

## Figure 2 - all definitions, x=IMD plot
## Prep dataframe
```{r}
df_plot_IMD <- table_cmm_join %>%
  pivot_longer(cols = c("Multimorbidity 2+", "Multimorbidity 3+", "Multimorbidity 3+ from 3+", "Mental-physical multimorbidity"), names_to = "definition", values_to = "count") %>%
  replace(is.na(.),0) %>%
  separate("age_gender_IMD", c("age", "gender", "IMD Decile")) %>%
  select(-gender) %>%
  mutate(age = case_when(
   age == '00to04' ~ "00 - 04", 
   age == '05to09' ~ "05 - 09",
   age == '10to14' ~ "10 - 14", 
   age == '15to19' ~ "15 - 19",
   age == '20to24' ~ "20 - 24", 
   age == '25to29' ~ "25 - 29",
   age == '30to34' ~ "30 - 34", 
   age == '35to39' ~ "35 - 39",
   age == '40to44' ~ "40 - 44", 
   age == '45to49' ~ "45 - 49",
   age == '50to54' ~ "50 - 54", 
   age == '55to59' ~ "55 - 59",
   age == '60to64' ~ "60 - 64", 
   age == '65to69' ~ "65 - 69",
   age == '70to74' ~ "70 - 74", 
   age == '75to79' ~ "75 - 79",
   age == '80to84' ~ "80 - 84", 
   age == '85to89' ~ "85 - 89",
   age == 'Over90' ~ "90+", 
)) %>%
  mutate(`IMD Decile` = case_when(
    `IMD Decile` == '01' ~ "1 (least deprived)",
    `IMD Decile` == '02' ~ "2",
    `IMD Decile` == '03' ~ "3",
    `IMD Decile` == '04' ~ "4",
    `IMD Decile` == '05' ~ "5",
    `IMD Decile` == '06' ~ "6",
    `IMD Decile` == '07' ~ "7",
    `IMD Decile` == '08' ~ "8",
    `IMD Decile` == '09' ~ "9",
    `IMD Decile` == '10' ~ "10 (most deprived)"
  )) %>%
  group_by(age, `IMD Decile`, definition) %>%
  summarise_if(is.numeric, sum) %>%
  mutate(prop = count/n *100)

df_plot_IMD$definition <- factor(df_plot_IMD$definition, levels = c("Multimorbidity 2+", "Multimorbidity 3+", "Multimorbidity 3+ from 3+", "Mental-physical multimorbidity"))

df_plot_IMD$`IMD Decile` <- factor(df_plot_IMD$`IMD Decile`, levels = c("1 (least deprived)", "2", "3", "4", "5", "6", "7", "8", "9", "10 (most deprived)"))
```

## Supp Figure 2 - all definitions, x=IMD plot
```{r}
df_plot_IMD %>% 
  filter(definition == "Multimorbidity 2+") %>%
  ggplot() +
  aes(age, prop, group = `IMD Decile`, colour = `IMD Decile`)+
  geom_line(size = 1)+
  theme(axis.ticks.x = element_line()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = c(.1, .87)) +
  #theme(legend.title=element_blank()) +
  scale_y_continuous(limits= c(0, 100), expand = c(0,0), 
                     breaks = scales::pretty_breaks(n = 4))+
  scale_x_discrete(expand = c(0,0))+
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "Age group (years)",
       y = "% population with multimorbidity 2+") +
  scale_colour_ptol()
  
ggsave("suppfig2_IMD_mm2.pdf", width = 16, height = 12)


df_plot_IMD %>% 
  filter(definition == "Multimorbidity 3+") %>%
  ggplot() +
  aes(age, prop, group = `IMD Decile`, colour = `IMD Decile`)+
  geom_line(size = 1)+
  theme(axis.ticks.x = element_line()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = c(.1, .87)) +
  #theme(legend.title=element_blank()) +
  scale_y_continuous(limits= c(0, 100), expand = c(0,0), 
                     breaks = scales::pretty_breaks(n = 4))+
  scale_x_discrete(expand = c(0,0))+
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "Age group (years)",
       y = "% population with multimorbidity 3+") +
  scale_colour_ptol()
  
ggsave("suppfig2_IMD_mm3.pdf", width = 16, height = 12)

df_plot_IMD %>% 
  filter(definition == "Multimorbidity 3+ from 3+") %>%
  ggplot() +
  aes(age, prop, group = `IMD Decile`, colour = `IMD Decile`)+
  geom_line(size = 1)+
  theme(axis.ticks.x = element_line()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = c(.1, .87)) +
  #theme(legend.title=element_blank()) +
  scale_y_continuous(limits= c(0, 100), expand = c(0,0), 
                     breaks = scales::pretty_breaks(n = 4))+
  scale_x_discrete(expand = c(0,0))+
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "Age group (years)",
       y = "% population with multimorbidity 3+ from 3+") +
  scale_colour_ptol()
  
ggsave("suppfig2_IMD_mm3from3.pdf", width = 16, height = 12)

df_plot_IMD %>% 
  filter(definition == "Mental-physical multimorbidity") %>%
  ggplot() +
  aes(age, prop, group = `IMD Decile`, colour = `IMD Decile`)+
  geom_line(size = 1)+
  theme(axis.ticks.x = element_line()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = c(.1, .87)) +
  #theme(legend.title=element_blank()) +
  scale_y_continuous(limits= c(0, 100), expand = c(0,0), 
                     breaks = scales::pretty_breaks(n = 4))+
  scale_x_discrete(expand = c(0,0))+
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "Age group (years)",
       y = "% population with mental-physical multimorbidity") +
  scale_colour_ptol()
  
ggsave("suppfig2_IMD_mp1.pdf", width = 16, height = 12)
```

## Figure 2 - all definitions, x=IMD plot
```{r}
df_plot_IMD <- df_plot_IMD %>%
  filter(`IMD Decile` %in% c("1 (least deprived)", "10 (most deprived)")) %>%
  rowwise %>%
  mutate(prop = ((prop.test(count, n, conf.level=0.95)$estimate)*100),
         lci = ((prop.test(count, n, conf.level=0.95)$conf.int[1])*100),
         uci = ((prop.test(count, n, conf.level=0.95)$conf.int[2])*100)) %>%
  mutate_if(is.numeric,
            round,
            digits = 2)

df_plot_IMD %>% 
  ggplot() +
  aes(age, prop, group = `IMD Decile`, colour = `IMD Decile`)+
  geom_line(size = 1)+
  geom_errorbar(aes(ymin=lci, ymax=uci), width=.1) +
  facet_wrap(~definition, scales="free")+
  theme(axis.ticks.x = element_line()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = c(.1, .87)) +
  scale_y_continuous(limits= c(0, 100), expand = c(0,0), 
                     breaks = scales::pretty_breaks(n = 4))+
  scale_x_discrete(expand = c(0,0))+
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "Age group (years)",
       y = "% population") +
  scale_colour_ptol()
  
ggsave("fig2_definitions_IMD.png", width = 16, height = 12)
ggsave("fig2_definitions_IMD.pdf", width = 16, height = 12)
```


## Supp figure 3 - mental-physical multimorbidity
## Prep dataframe
```{r}
colnames(table_mp_join) <- c("age_gender_IMD", "Depression anxiety physical", "Dementia physical", "Substance alcohol misuse physical", "Schizophrenia bipolar physical", "n")

df_plot_mp_IMD <- table_mp_join %>%
  pivot_longer(cols = c("Depression anxiety physical", "Dementia physical", "Substance alcohol misuse physical", "Schizophrenia bipolar physical"), names_to = "definition", values_to = "count") %>%
  replace(is.na(.),0) %>%
  separate("age_gender_IMD", c("age", "gender", "IMD Decile")) %>%
  select(-gender) %>%
  group_by(age, `IMD Decile`, definition) %>%
  summarise_if(is.numeric, sum) %>%
 mutate(age = case_when(
   age == '00to04' ~ "00 - 04", 
   age == '05to09' ~ "05 - 09",
   age == '10to14' ~ "10 - 14", 
   age == '15to19' ~ "15 - 19",
   age == '20to24' ~ "20 - 24", 
   age == '25to29' ~ "25 - 29",
   age == '30to34' ~ "30 - 34", 
   age == '35to39' ~ "35 - 39",
   age == '40to44' ~ "40 - 44", 
   age == '45to49' ~ "45 - 49",
   age == '50to54' ~ "50 - 54", 
   age == '55to59' ~ "55 - 59",
   age == '60to64' ~ "60 - 64", 
   age == '65to69' ~ "65 - 69",
   age == '70to74' ~ "70 - 74", 
   age == '75to79' ~ "75 - 79",
   age == '80to84' ~ "80 - 84", 
   age == '85to89' ~ "85 - 89",
   age == 'Over90' ~ "90+", 
  )) %>%
  mutate(`IMD Decile` = case_when(
    `IMD Decile` == '01' ~ "1 (least deprived)",
    `IMD Decile` == '02' ~ "2",
    `IMD Decile` == '03' ~ "3",
    `IMD Decile` == '04' ~ "4",
    `IMD Decile` == '05' ~ "5",
    `IMD Decile` == '06' ~ "6",
    `IMD Decile` == '07' ~ "7",
    `IMD Decile` == '08' ~ "8",
    `IMD Decile` == '09' ~ "9",
    `IMD Decile` == '10' ~ "10 (most deprived)"
  )) %>%
  group_by(age, `IMD Decile`, definition) %>%
  summarise_if(is.numeric, sum) %>%
  mutate(prop = count/n *100)

df_plot_mp_IMD$definition <- factor(df_plot_mp_IMD$definition, levels = c("Depression anxiety physical", "Substance alcohol misuse physical", "Schizophrenia bipolar physical", "Dementia physical"))

df_plot_mp_IMD$`IMD Decile` <- factor(df_plot_mp_IMD$`IMD Decile`, levels = c("1 (least deprived)", "2", "3", "4", "5", "6", "7", "8", "9", "10 (most deprived)"))
```


## Supp Figures 5-8 - all definitions, x=IMD plot
```{r}
df_plot_mp_IMD %>% 
  filter(definition == "Depression anxiety physical") %>%
  ggplot() +
  aes(age, prop, group = `IMD Decile`, colour = `IMD Decile`)+
  geom_line(size = 1)+
  theme(axis.ticks.x = element_line()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = c(.1, .87)) +
  #theme(legend.title=element_blank()) +
  scale_y_continuous(limits= c(0, 45), expand = c(0,0), 
                     breaks = scales::pretty_breaks(n = 4))+
  scale_x_discrete(expand = c(0,0))+
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "Age group (years)",
       y = "% population with depression/anxiety-physical multiorbidity") +
  scale_colour_ptol()
  
ggsave("suppfig3_IMD_depanx.pdf", width = 16, height = 12)


df_plot_mp_IMD %>% 
  filter(definition == "Substance alcohol misuse physical") %>%
  ggplot() +
  aes(age, prop, group = `IMD Decile`, colour = `IMD Decile`)+
  geom_line(size = 1)+
  theme(axis.ticks.x = element_line()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = c(.1, .87)) +
  #theme(legend.title=element_blank()) +
  scale_y_continuous(limits= c(0, 45), expand = c(0,0), 
                     breaks = scales::pretty_breaks(n = 4))+
  scale_x_discrete(expand = c(0,0))+
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "Age group (years)",
       y = "% population with substance/alcohol-physical multimorbidity") +
  scale_colour_ptol()
  
ggsave("suppfig3_IMD_subst.pdf", width = 16, height = 12)

df_plot_mp_IMD %>% 
  filter(definition == "Schizophrenia bipolar physical") %>%
  ggplot() +
  aes(age, prop, group = `IMD Decile`, colour = `IMD Decile`)+
  geom_line(size = 1)+
  theme(axis.ticks.x = element_line()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = c(.1, .87)) +
  #theme(legend.title=element_blank()) +
  scale_y_continuous(limits= c(0, 45), expand = c(0,0), 
                     breaks = scales::pretty_breaks(n = 4))+
  scale_x_discrete(expand = c(0,0))+
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "Age group (years)",
       y = "% population with schizophrenia/bipolar-physical multimorbidity") +
  scale_colour_ptol()
  
ggsave("suppfig3_IMD_schiz.pdf", width = 16, height = 12)

df_plot_mp_IMD %>% 
  filter(definition == "Substance alcohol misuse physical") %>%
  ggplot() +
  aes(age, prop, group = `IMD Decile`, colour = `IMD Decile`)+
  geom_line(size = 1)+
  theme(axis.ticks.x = element_line()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = c(.1, .87)) +
  #theme(legend.title=element_blank()) +
  scale_y_continuous(limits= c(0, 45), expand = c(0,0), 
                     breaks = scales::pretty_breaks(n = 4))+
  scale_x_discrete(expand = c(0,0))+
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "Age group (years)",
       y = "% population with demential-physical multimorbidity") +
  scale_colour_ptol()
  
ggsave("suppfig3_IMD_dementia.pdf", width = 16, height = 12)
```

## Figure 3 - Mental-physical multimorbidity subdivisions
```{r}
df_plot_mp_IMD <- df_plot_mp_IMD %>% 
  filter(`IMD Decile` %in% c("1 (least deprived)", "10 (most deprived)")) %>%
  rowwise %>%
  mutate(prop = ((prop.test(count, n, conf.level=0.95)$estimate)*100),
         lci = ((prop.test(count, n, conf.level=0.95)$conf.int[1])*100),
         uci = ((prop.test(count, n, conf.level=0.95)$conf.int[2])*100)) %>%
  mutate_if(is.numeric,
            round,
            digits = 2)

df_plot_mp_IMD %>%
  ggplot() +
  aes(age, prop, group = `IMD Decile`, colour = `IMD Decile`)+
  geom_line(size = 2)+
  geom_errorbar(aes(ymin=lci, ymax=uci), width=.1) +
  facet_wrap(~definition, scales="free")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(legend.position = c(.1, .87)) +
  #theme(legend.title=element_blank()) +
  scale_y_continuous(limits= c(0, 45), expand = c(0,0), 
                     breaks = scales::pretty_breaks(n = 4))+
  scale_x_discrete(expand = c(0,0))+
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "Age group (years)",
       y = "% population") +
  scale_colour_ptol()
  
ggsave("suppfig3_mpdefinitions_IMD.pdf", width = 16, height = 12)
```
